On the presentation of orthogonal groups over polynomial rings

### 2019-08-14 dl-paper.tex
Minor

\section{Proof of the main result}
\subsection{Presentation of Steinberg groups by homogeneous generators}
\label{sec:presentation}
Let $M\trianglelefteq A$ be an ideal of a commutative ring $A$.
We consider $A[t, t\inv]$ as a $\mathbb{Z}$-graded ring in which $t$ has degree $1$.
We denote by $B = B(A, M)$ the subring $A[t] + M[t\inv] \subseteq A[t, t\inv]$ with the induced grading.
As an $A$-module $B$ decomposes as $\oplus_{d\in\mathbb Z}B_d$ where $B_d=M \cdot t^d$ for $d<0$, and $B_d=A \cdot t^d$ for $d\geq0$. Obviously, $B = A[t]$ in the case $M=0$ and $B = A[t, t\inv]$ in the case $M=A$.

@@

\begin{prop} \label{lemma33} For $\Phi=\rA_{\geq 4}, \rD_{\geq 5}, \rE_{6,7,8}$ and $n \geq 1$ the map $i_n\colon \St^{\leq n}(\Phi, B) \to \St^{\leq n+1}(\Phi, B)$ induced by the natural embedding of generators, is an isomorphism. In particular, the obvious map $\St^{\leq 1}(\Phi, B) \to \St(\Phi, B)$ is an isomorphism. \end{prop}
\begin{proof}
 We need to construct a map $j_n$ which would be the inverse of $i_n$. 
 We start with a word map $\widetilde{j}_n^\Phi \colon F\langle \mathcal{X}^\Phi_{\leq n+1} \rangle \to \St^{\leq n}(\Phi, B)$ defined via
 \[ \widetilde{j}_n(x_\alpha(at^k)) = \begin{cases} x_\alpha(at^k), & k\leq n; \\
      [x_{\alpha - \beta} (N_{\alpha-\beta, \beta} \cdot at^{k-1}), x_{\beta}(t)], & k = n+1, \end{cases} \]
 where $\beta$ is an arbitrary root of $\Phi$ forming a sharp angle with $\alpha$.
 Standard argument (cf.~\cite[Proposition~1.1]{Re75} or~\cite[Proposition~3.2.2]{RS76}) shows that $\widetilde{j}_n$ does not depend on the choice of $\beta$.

@@

\subsection{The subgroups \texorpdfstring{$P_\alpha(0)$}{Pa(0)}, \texorpdfstring{$P_\alpha(*)$}{Pa(*)} and their properties.} 
For a root system $\Phi$ consider the following subsets of $\Phi$:
\begin{align} Z_+(\alpha) & = \{ \beta \in \Phi \mid \langle \alpha, \beta \rangle > 0 \}, \\
   Z_0(\alpha) & = \{ \beta \in \Phi \mid \alpha + \beta \not\in \Phi,\ \langle \alpha, \beta \rangle = 0 \}, \\
   Z(\alpha)   & = Z_0(\alpha) \sqcup Z_+(\alpha). \end{align} 
Clearly, $Z_0(\alpha)$ (resp. $Z_+(\alpha)$) is a reductive (resp. special) subset of $\Phi$.

@@

\subsection{The subgroups \texorpdfstring{$K(\alpha, \beta)$}{K(a,b)} and their properties}
Throughout this subsection $\alpha, \beta$ denote a fixed pair of roots of $\Phi$ forming a sharp angle (i.\,e. $\langle \alpha, \beta \rangle = 1$). We denote by $\Psi$ the subsystem of type $\rA_2$ generated by $\alpha$ and $\beta$.
\begin{df} Denote by $K(\alpha, \beta)$ the subgroup of $\St(\Phi, A[X, X^{-1}])$ generated by the following elements (as before, $\xi \in A[X]$, $f\in M[X]$):
 \begin{enumerate}
  \item $z_\gamma(Xf, \xi)$, for all $\gamma \in \Phi \setminus \Psi$;
  \item $x_{-\alpha}(X^2f)$, $x_{-\beta}(X^2f)$;
  \item $x_{\alpha}(Xf)$, $x_\beta(Xf)$;
  \item $x_{\alpha-\beta}(Xf)$, $x_{\beta-\alpha}(Xf)$;
  \item $c_{\delta}(f, X\xi)$ for some fixed root $\delta \in \Phi \setminus \Psi$. \end{enumerate} \end{df}

@@

For shortness we denote the subgroup $\overline{\St}(\Phi, A, M)$ by $G_M^0$.
Notice that $G_M^0$ can be considered as a subgroup of $\St(\Phi, R)$.
\[ \xymatrix{ G_M^0 \ar@{^{(}->}[r]^{i_+} \ar@{^{(}->}[d]^{i_-} & \overline{\St}(\Phi, A[X], M[X]) \ar^{j_+^M}[d] \ar@{^{(}->}[r] & \St(\Phi, A[X]) \ar[dl]^{j_+} \\
              \St(\Phi, A[X\inv]) \ar[r]^{j_-} & \St(\Phi, R) &  } \] 
We denote by $\overline{G}^{\geq 0}_M$ the image of the map $j_+^M$.

Denote by $V$ the subset of \[\St(\Phi, R) \times \St(\Phi, A[X\inv]) \times (1 + M)^\times\] consisting of those triples $(g, h, u)$ for which
\[p(g, h, u) := g \cdot j_-(h) \cdot \{ X, u \}\] belongs to $\overline{G}_M^{\geq 0}$.

@@

The isomorphism between the sets $\overline{V}$ and $\overline{V}_T$ is given by the following two maps, which are easily seen to be mutually inverse to each other:
\begin{equation} \label{eq:VVT} \xymatrix @R0.3pc {\overline{V} \ar[r]^{\cong} & \ar[l] \overline{V}_T \\ (g, [h], u) \ar@{|->}[r] & [p(g, h, u), h^{-1}, u] \\ (p \cdot j_-(h) \cdot \{u, X\}, [h^{-1}], u) & \ar@{|->}[l] [p, h, u]. } \end{equation}
The above isomorphism allows us to treat $\overline{V}$ and $\overline{V}_T$ as a single object, for which we can interchangeably use either of the two notations,
 depending on which of them is more convenient in a given situation.
For example, specifying the action of $\St(\Phi, B)$ in terms of $\overline{V}$ leads to much shorter calculations in Lemmas~\ref{R3_leq0_1}--\ref{R2_0_1},
 while the statements of~\cref{prop43} and~\cref{lem:action} are more natural when expressed in terms of~$\overline{V}_T$.

Now we are ready to proceed with the construction of the action of $\St(\Phi, B)$ on $\overline{V}$.

We start by defining for $\alpha \in \Phi$, $a \in A$ a partial function $t_\alpha(aX^{-1}) \colon V \not\to V$.
This function is defined for the triples $(g, h, u)$ satisfying $p(g, h, u) \in P_\alpha(*) \subseteq \overline{G}_M^{\geq 0}$.
If $p(g, h, u)$ belongs to $P_\alpha(m)$ for some $m \in M$, then $t_\alpha(aX^{-1})$ is defined via the following identity:
\begin{equation} \label{T_1} t_\alpha(aX^{-1}) (g, h, u) = \left( x_\alpha(aX^{-1})\cdot g ,\ h \cdot x_\alpha\left(-\tfrac{aX^{-1}}{1 + am}\right),\ u \cdot (1 + am)\right).\end{equation}

>>>>

\section{Proof of the main result}
The main result of this section is~\cref{thm:P1glueing}, which is a direct generalization of~\cite[Proposition~4.3]{Tu83}. The proof of this theorem is rather long and technical, so before setting out technical details, we briefly describe the main steps of this proof.

The object playing a key role in the proof of~\cref{thm:P1glueing} is a certain action of the group $G = \St(\Phi, A[X\inv] + M[X])$ on a certain set $\overline{V}$. To be able to write an explicit formula for this action we need two major ingredients. The first one is a presentation of $G$, which involves much fewer generations and relations. We obtain this presentation in~\cref{lemma33}. The other ingredient are the subgroups $P_\alpha(0)$ of $\St(\Phi, A[X, X\inv])$, whose definition and properties are given in Sections~\ref{sec:Pa0-basic}--\ref{sec:Pa0-Steinberg}.

\subsection{Presentation of Steinberg groups by homogeneous generators}
\label{sec:presentation}
Let $M\trianglelefteq A$ be an ideal of a commutative ring $A$.
We consider $A[t, t\inv]$ as a $\mathbb{Z}$-graded ring in which $t$ has degree $1$.
We denote by $B = B(A, M)$ the subring $A[t] + M[t\inv] \subseteq A[t, t\inv]$ with the induced grading.
As an $A$-module $B$ decomposes as $\oplus_{d\in\mathbb Z}B_d$ where $B_d=M \cdot t^d$ for $d<0$, and $B_d=A \cdot t^d$ for $d\geq0$. Obviously, $B = A[t]$ in the case $M=0$ and $B = A[t, t\inv]$ in the case $M=A$.

@@

The following proposition is the main result of this subsection and also a direct generalization of~\cite[Lemma~3.3]{Tu83}.
\begin{prop} \label{lemma33} For $\Phi=\rA_{\geq 4}, \rD_{\geq 5}, \rE_{6,7,8}$ and $n \geq 1$ the map $i_n\colon \St^{\leq n}(\Phi, B) \to \St^{\leq n+1}(\Phi, B)$ induced by the natural embedding of generators, is an isomorphism. In particular, the obvious map $\St^{\leq 1}(\Phi, B) \to \St(\Phi, B)$ is an isomorphism. \end{prop}
\begin{proof}
 We need to construct a map $j_n$ which would be the inverse of $i_n$. 
 We start with a word map $\widetilde{j}_n^\Phi \colon F\langle \mathcal{X}^\Phi_{\leq n+1} \rangle \to \St^{\leq n}(\Phi, B)$ defined via
 \[ \widetilde{j}_n(x_\alpha(at^k)) = \begin{cases} x_\alpha(at^k), & k\leq n; \\
      [x_{\alpha - \beta} (N_{\alpha-\beta, \beta} \cdot at^{k-1}), x_{\beta}(t)], & k = n+1, \end{cases} \]
 where $\beta$ is an arbitrary root of $\Phi$ forming a sharp angle with $\alpha$.
 Standard argument (cf.~\cite[Proposition~1.1]{Re75} or~\cite[Proposition~3.2.2]{RS76}) shows that $\widetilde{j}_n$ does not depend on the choice of $\beta$.

@@

\subsection{The subgroups \texorpdfstring{$P_\alpha(0)$}{Pa(0)}, \texorpdfstring{$P_\alpha(*)$}{Pa(*)} and their properties.} \label{sec:Pa0-basic}
For a root system $\Phi$ consider the following subsets of $\Phi$:
\begin{align} Z_+(\alpha) & = \{ \beta \in \Phi \mid \langle \alpha, \beta \rangle > 0 \}, \\
   Z_0(\alpha) & = \{ \beta \in \Phi \mid \alpha + \beta \not\in \Phi,\ \langle \alpha, \beta \rangle = 0 \}, \\
   Z(\alpha)   & = Z_0(\alpha) \sqcup Z_+(\alpha). \end{align} 
Clearly, $Z_0(\alpha)$ (resp. $Z_+(\alpha)$) is a reductive (resp. special) subset of $\Phi$.

@@

\subsection{The subgroups \texorpdfstring{$K(\alpha, \beta)$}{K(a,b)} and their properties} \label{sec:Pa0-Steinberg}
Throughout this subsection $\alpha, \beta$ denote a fixed pair of roots of $\Phi$ forming a sharp angle (i.\,e. $\langle \alpha, \beta \rangle = 1$). We denote by $\Psi$ the subsystem of type $\rA_2$ generated by $\alpha$ and $\beta$.
\begin{df} Denote by $K(\alpha, \beta)$ the subgroup of $\St(\Phi, A[X, X^{-1}])$ generated by the following elements (as before, $\xi \in A[X]$, $f\in M[X]$):
 \begin{enumerate}
  \item $z_\gamma(Xf, \xi)$, for all $\gamma \in \Phi \setminus \Psi$;
  \item $x_{-\alpha}(X^2f)$, $x_{-\beta}(X^2f)$;
  \item $x_{\alpha}(Xf)$, $x_\beta(Xf)$;
  \item $x_{\alpha-\beta}(Xf)$, $x_{\beta-\alpha}(Xf)$;
  \item $c_{\delta}(f, X\xi)$ for some fixed root $\delta \in \Phi \setminus \Psi$. \end{enumerate} \end{df}

@@

For shortness we denote the subgroup $\overline{\St}(\Phi, A, M)$ by $G_M^0$.
Notice that $G_M^0$ can be considered as a subgroup of $\St(\Phi, R)$.
\[ \xymatrix{ G_M^0 \ar@{^{(}->}[r]^{i_+} \ar@{^{(}->}[d]^{i_-} & \overline{\St}(\Phi, A[X], M[X]) \ar^{j_+^M}[d] \ar@{^{(}->}[r] & \St(\Phi, A[X]) \ar[dl]^{j_+} \\
              \St(\Phi, A[X\inv]) \ar[r]^{j_-} & \St(\Phi, R) &  } \] 
We denote by $\overline{G}^{\geq 0}_M$ the image of the map $j_+^M$. 
Denote by $V$ the subset of \[\St(\Phi, R) \times \St(\Phi, A[X\inv]) \times (1 + M)^\times\] consisting of those triples $(g, h, u)$ for which
\[p(g, h, u) := g \cdot j_-(h) \cdot \{ X, u \}\] belongs to $\overline{G}_M^{\geq 0}$.

@@

The isomorphism between the sets $\overline{V}$ and $\overline{V}_T$ is given by the following two maps, which are easily seen to be mutually inverse to each other:
\begin{equation} \label{eq:VVT} \xymatrix @R0.3pc {\overline{V} \ar[r]^{\cong} & \ar[l] \overline{V}_T \\ (g, [h], u) \ar@{|->}[r] & [p(g, h, u), h^{-1}, u] \\ (p \cdot j_-(h) \cdot \{u, X\}, [h^{-1}], u) & \ar@{|->}[l] [p, h, u]. } \end{equation}
The above isomorphism allows us to treat $\overline{V}$ and $\overline{V}_T$ as a single object, for which we can interchangeably use either of the two notations,
 depending on which of them is more convenient in a given situation.
For example, specifying the action of $\St(\Phi, B)$ in terms of $\overline{V}$ leads to much shorter calculations in Lemmas~\ref{R3_leq0_1}--\ref{R2_0_1},
 while the statements of~\cref{prop43} and~\cref{lem:action} look more natural when formulated in terms of~$\overline{V}_T$.

Now we are ready to proceed with the construction of the action of $\St(\Phi, B)$ on $\overline{V}$. We start by defining for $\alpha \in \Phi$, $a \in A$ a partial function $t_\alpha(aX^{-1}) \colon V \not\to V$.
This function is defined for the triples $(g, h, u)$ satisfying $p(g, h, u) \in P_\alpha(*) \subseteq \overline{G}_M^{\geq 0}$.
If $p(g, h, u)$ belongs to $P_\alpha(m)$ for some $m \in M$, then $t_\alpha(aX^{-1})$ is defined via the following identity:
\begin{equation} \label{T_1} t_\alpha(aX^{-1}) (g, h, u) = \left( x_\alpha(aX^{-1})\cdot g ,\ h \cdot x_\alpha\left(-\tfrac{aX^{-1}}{1 + am}\right),\ u \cdot (1 + am)\right).\end{equation}
